---
title: "Homework_04"
author: "Kelsey Soule"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. 

Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

a. $x^{a^{b}}$  
b. $(x^a)^b$  
c. $3x^3+2x^2+1$  

```{r}
x <- 1.1
a <- 2.2
b <- 3.3

z <- x^{a^{b}}
print(z)

z <- (x^a)^b
print(z)

z <- 3*(x^3)+2*(x^2)+1
print(z)

```

### 2.

Using the rep and seq functions, create the following vectors:

a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
a <- c(seq(from=1,to=8),seq(from=7,to=1))
print(a)

vect <- seq(from=1,to=5)
b <- rep(vect,times=vect)
print(b)

vect_2 <- seq(from=5,to=1)
c <- rep(vect_2,times=vect)
print(c)
```

### 3. 
Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (`asin()`, `acos()`, or `atan()`), convert these numbers into polar coordinates.

```{r}
coord <- runif(2,min = 0,max = 10)
print(coord)

r <- sqrt((coord[1])^2 + (coord[2])^2)

theta <- atan((coord[2])/(coord[1]))
polar <- c(r,theta)
print(polar)
```

### 4. 
Create a vector queue <- c("sheep", "fox", "owl", "ant") where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update queue as:

a. the serpent arrives and gets in line;
b. the sheep enters the ark;
c. the donkey arrives and talks his way to the front of the line;
d. the serpent gets impatient and leaves;
e. the owl gets bored and leaves;
f. the aphid arrives and the ant invites him to cut in line.
g. Finally, determine the position of the aphid in the line.

```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)

queue <- c(queue,"serpent")
print(queue)

queue <- setdiff(queue,"sheep")
print(queue)

queue <- c("donkey",queue)
print(queue)

queue <- setdiff(queue,"serpent")
print(queue)

queue <- setdiff(queue,"owl")
print(queue)

queue <- c(queue[1:2],"aphid",queue[3])
print(queue)

which(queue=="aphid")
```

