---
title: "Homework_05"
author: "Kelsey Soule"
date: "2024-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

### 1. 

Assign to the variable n_dims a single random integer between 3 and 10.  

* Create a vector of consecutive integers from 1 to n_dims^2.  
* Use the sample function to randomly reshuffle these values.  
* create a square matrix with these elements.  
* print out the matrix.  
* find a function in r to transpose the matrix.  
* print it out again and note how it has changed.  
* calculate the sum and the mean of the elements in the first row and then the last row.  
* read about the eigen() function and use it on your matrix.  
* look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?  
* dig in with the typeof() function to figure out their type.  
* if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.  
```{r}
n_dims <- sample(3:10,1)

vec <- c(1:(n_dims)^2)

sample(vec)

matr <- matrix(data = vec, nrow = n_dims)

print(matr)

matr <- t(matr)

print(matr)

sum(matr[1,])
mean(matr[1,])
sum(matr[n_dims,])
mean(matr[n_dims,])

eigen(matr)
```
The `$values` are eigenvalues of the matrix which are the characteristic roots of the matrix, and they are scalar. The `$vectors` are the eigenvectors of the matrix, and are vectors so they are not scalar and have direction.
```{r}
typeof(eigen(matr)$values)
typeof(eigen(matr)$vectors)
# these are complex numbers
```

### 2. 
Create a list with the following named elements:

* my_matrix, which is a 4 x 4 matrix filled with random uniform values
* my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.  
* my_letters, which is a 26-element vector of all the lower-case letters in random order.  

Then, complete the following steps:

* create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.  
* use the typeof() function to confirm the underlying data types of each component in this list
* combine the underlying elements from the new list into a single atomic vector with the c() function.
* what is the data type of this vector?

```{r}
mylist <- list(my_matrix=matrix(data = runif(16),nrow = 4), my_logical=(runif(100))<0.5, my_letters=sample(letters))
print(mylist)

newlist <- list(mylist$my_matrix[2,2], mylist$my_logical[2], mylist$my_letters[2])
print(newlist) 

typeof(newlist[[1]])
typeof(newlist[[2]])
typeof(newlist[[3]])

myvec <- c(newlist[[1]],newlist[[2]],newlist[[3]])
typeof(myvec)
```

### 3. 
Create a data frame with the two variables (= columns) and 26 cases (= rows) below:

* call the first variable my_unis and fill it with 26 random uniform values from 0 to 10  
* call the second variable my_letters and fill it with 26 capital letters in random order.  
* for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.  
* for the first variable, write a single line of R code to identify which rows have the missing values.  
* re-order the entire data frame to arrange the second variable in alphabetical order. 
* calculate the column mean for the first variable.

```{r}
myframe <- data.frame(my_unis=runif(26,min = 0,max = 10),my_letters=sample(LETTERS))
print(myframe)

myframe[sample(1:26,4),1]<-NA

which(is.na(myframe$my_unis))

arrange(myframe,my_letters)

mean(myframe$my_unis,na.rm = TRUE)

```

